Reproducible Research Week 2 Project
=========================================
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load and Process Data

```{r}
library(tidyverse)
df <- read.csv("activity.csv", sep = ",", header = TRUE)
df$date <- as.Date(as.character(df$date, "%Y%m%d"))
```


#Total Number of Steps Taken each day
#Histogram of the total number of steps taken each day.
```{r}
df[!is.na(df$steps) == TRUE,] %>% 
  group_by(date) %>% 
  summarise(total_steps = sum(steps)) %>% 
  ggplot() +
  geom_histogram(aes(x=total_steps), binwidth = 1000)
```

#Mean and Median of steps taken per day
```{r}
df_date <- df[!is.na(df$steps) == TRUE,] %>% 
  group_by(date) %>% 
  summarise(total_steps = sum(steps))  
mean(df_date$total_steps)
median(df_date$total_steps)
```

#Time series plot of the average number of steps taken
```{r}
x <- unique(df$interval)
y <- tapply(df$steps, df$interval, mean, na.rm=T)
plot(x, y, type = "l",xlab="Interval", ylab="Steps", main = "Average Number of Steps taken in 5-min Intervals")
```

#The 5-minute interval that, on average, contains the maximum number of steps
```{r}
df_int <- df[!is.na(df$steps) == TRUE,] %>% 
  group_by(interval) %>% 
  summarise(steps =mean(steps))
df_int[df_int$steps == max(df_int$steps),]
```

#Number of Rows with Missing Values
```{r}
nrow(df[is.na(df$steps) == TRUE,])
```

#Impute Missing Values
```{r}
df2 <- df
int_avg <- as.data.frame(tapply(df2$steps, df2$interval, mean, na.rm=T))
int_avg$interval <- rownames(int_avg)
int_avg$interval <- as.integer(int_avg$interval)
names(int_avg) <- c("avg","interval")
df2 <- left_join(df2,int_avg, by = "interval")
df2$steps <- ifelse(is.na(df2$steps) == TRUE, df2$avg, df2$steps)
df2 <- df2[ ,-c(4)]
```

#Histogram of the total number of steps taken each day after missing values are imputed 
```{r}
  df2%>% 
  group_by(date) %>% 
  summarise(total_steps = sum(steps)) %>% 
  ggplot() +
  geom_histogram(aes(x=total_steps), binwidth = 1000)
```

#Mean and Median of Total Steps by date
```{r}
df2_date <- df2%>% 
            group_by(date) %>% 
            summarise(total_steps = sum(steps)) 
mean(df2_date$total_steps)
median(df2_date$total_steps)
```

#Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r}
df2$weekday_end <- ifelse(weekdays(df2$date) %in% c("Saturday","Sunday"), "weekend", "weekday")
df2%>% 
  group_by(interval, weekday_end) %>% 
  summarise(avg_steps = mean(steps)) %>% 
  ggplot() +
  geom_line(aes(x=interval, y=avg_steps)) +
  facet_grid(weekday_end ~.) +
  labs(x="Interval", y="Number of Steps", title="Comparison - Average Number of Steps in each Interval")
```





